---
output: 
  bookdown::pdf_document2:
    number_sections: FALSE
    toc: FALSE
urlcolor: blue
fontsize: 12pt
header-includes:
    - \usepackage{caption}
    - \usepackage{booktabs}
    - \usepackage{mathpazo}
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{fancyhdr}
    - \pagestyle{fancy}
    - \setlength{\headheight}{16pt}
    - \captionsetup{labelformat=empty, font=singlespacing}
    - \fancyhead[L]{\selectfont Moving across gradients.}
bibliography: refs.bib
csl: amnat.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    warning = FALSE,
    message = FALSE,
    eval = TRUE
)
knitr::opts_knit$set(root.dir = '../')
```

```{r prep, results = 'hide'}
library(data.table)
library(kableExtra)
library(lme4)

source('R/variables.R')

locs <- readRDS('data/derived-data/cleaned-locs.Rds')
m1 <- readRDS('data/derived-data/figure2-m1.Rds')
```

\thispagestyle{empty}

### **Supplement 3: Foraging landscape efficiency model results and sensitivity analysis**

The ecology of individual differences empirically applied to space-use and movement tactics

Quinn M. R. Webber$^{1,*}$, Michel P. Laforge$^{2,*,+}$, Maegwin Bonar$^{2}$, Alec L. Robitaille$^{2}$, Christopher Hart$^{2}$, Sana Zabihi-Seissan$^{2}$, Eric Vander Wal$^{1,2}$.


1. Cognitive and Behavioural Ecology Interdisciplinary Program, Memorial University of Newfoundland, St. John’s, Newfoundland, Canada

2. Department of Biology, Memorial University of Newfoundland, St. John’s, Newfoundland, Canada


$*$ These authors contributed equally to the preparation of this manuscript.


$+$ Corresponding author e-mail: m.laforge@mun.ca

The American Naturalist

\newpage
\setcounter{page}{1}

### **Supplement 3: Foraging landscape efficiency model results and sensitivity analysis**


(ref:tabs31) **Table S3.1**: Beta coefficients and 95% confidence interval (CI) for model of forage habitat selection efficiency for caribou in Newfoundland as a function of area-restricted search, range-use ratio and forage patch heterogeneity.

```{r tab1}
coefs <- fixef(m1)
SEs <- c(sqrt(diag(vcov(m1))))
LCI <- coefs - (SEs * 1.96)
UCI <- coefs + (SEs * 1.96)

h <- c('', '', '95 % CI' = 2)

names(coefs) <- c('Intercept', 'First passage time',
                  'Heterogeneity', 'Range-use ratio',
                  'First Passage time × heterogeneity',
                  'Heterogeneity × range-use ratio')

kable(data.table(Variable = names(coefs), 
           Estimate = coefs,
           Lower = LCI,
           Upper = UCI),
      digits = 3,
			align = "l",
			format = 'latex',
			booktabs = TRUE, 
			caption = '(ref:tabs31)',
			linesep = c('')) %>%
	add_header_above(h) %>%
  kable_styling(font_size = 10,
                latex_options = "hold_position")
```



![**Figure S3.1**: Sensitivity analysis for percentile of the landscape that is designated as high-quality patch for patch use models to predict the effect of first-passage time and forage patch aggregation (Moran’s I) and range-use ratio and forage patch aggregation on foraging efficiency for caribou (*Rangifer tarandus*, n = `r locs[, uniqueN(ANIMAL_ID)]`) in Newfoundland. Top two rows represent the response to first-passage time, bottom two rows represent response to range-use ratio (the other variable is held at its mean). Cell values represent the relative proportion of time an individual spends in high quality foraging habitat (the top x% of mean values of instantaneous rate of growth). Note, 20% of the landscape was designated as high-quality patch for all results displayed in the main text. ](../graphics/Supplement3/FigS3_1.png)






